---
title: "Route to Happiness: Unravelling the Link between Transportation and Quality of Life"
author: "Mekhala Kumar"
editor: visual
date: "2023-12-14"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(stargazer) 
library(skimr)
library(readxl)
library(zipcodeR)
library(ggplot2)
library(plotly)
library(MASS)
library(ggeffects)
library(brant)
library(performance)
library(descr)
library(gt)
library(formatR)

```

## Primary Research Question

a.  How do modes of transport impact satisfaction of life?
b.  How do modes of transport affect one's satisfaction with quality of life?

## Reading in data

Dataset- National Community Livability Survey

```{r}
publictransit<-read_excel("_data/Deliverable-3b-NCLS-Dataset.xlsx",sheet="Data")
backup<-publictransit
# Selecting the required variables
publictransit<-publictransit%>%dplyr::select(Hhveh,Walkrate,Transitrate,transitavail,transitavailns,LifeSat,Comqol,Hhincscale,AgeScale,Edscale,Male,White,Black,AIAN,Hawaiian,Asian,Otherrace,ZipCode,Male)
summary(publictransit)
#Assigning missing values and transforming variables wherever necessary
publictransit[publictransit=="."]<-NA
publictransit$lifesat<-as.numeric(publictransit$LifeSat)
publictransit$qol<-as.factor(publictransit$Comqol)
publictransit$hhveh<-as.numeric(publictransit$Hhveh)
publictransit$walkrate<-as.numeric(publictransit$Walkrate)
publictransit$transitrate<-as.numeric(publictransit$Transitrate)
publictransit$ptavail<-as.numeric(publictransit$transitavail)
publictransit$ptavailns<-as.numeric(publictransit$transitavailns)
publictransit$male<-as.factor(publictransit$Male)

publictransit<-publictransit%>%dplyr::select(-c(LifeSat,Comqol,Hhveh,Walkrate,Transitrate,transitavail,transitavailns,Male))
# Summary Statistics at a Glance
skim(publictransit)
publictransit <- publictransit[complete.cases(publictransit), ]

```

## Primary Explanatory Independent Variables

1.  Hhveh- How many working vehicles (cars, trucks, and motorcycles) are available in your household? This ranges from 0 to 3.
2.  Walkrate- Rate the quality of walkability/accessibility in your community (scale 1-5, very poor to very good).
3.  Transitqual- I created this variable based on transitavail (whether there is public transit) and transitrate (quality of public transit).The value=0, if there is no public transit and then for those who do have public transit in their community, the values range from 1-5.

```{r}
# Creating the perceived public transit quality variable based on transitavail and transitrate
publictransit<-publictransit%>%
  mutate(pt_qual=case_when(
    ptavail==0~0,
    ptavail==1 & transitrate==1~1,
    ptavail==1 & transitrate==2~2,
    ptavail==1 & transitrate==3~3,
    ptavail==1 & transitrate==4~4,
    ptavail==1 & transitrate==5~5
  ))
# Sanity check for new perceived public transit quality variable created
table(publictransit$ptavail)
publictransit%>%
  dplyr::select(ptavail,transitrate,pt_qual)%>%
  head()
typeof(publictransit$pt_qual)
# Converting public transit quality to factor
publictransit$ptqual<-as.factor(publictransit$pt_qual)
publictransit<-publictransit%>%dplyr::select(-c(pt_qual,ptavail,ptavailns,transitrate))

# Creating factor type variables for number of vehicles and walkability
publictransit$walk_fac<-as.factor(publictransit$walkrate)
publictransit$car_fac<-as.factor(publictransit$hhveh)
```

### For the control variable- geographic region

```{r}
# Taking data from the package zipcodeR
zip<-zip_code_db
# Selecting the required variables from the zipcode dataset
zip<-zip%>%dplyr::select(zipcode,population,population_density)
# Comparing the zipcode information in the 2 datasets
typeof(publictransit$zipcode)
typeof(zip$zipcode)
publictransit%>%distinct(ZipCode)
```

## Control Variables

1.  Geographic region- Urban/urban cluster/rural

2.  hhincscale- Income

3.  AgeScale- Age

4.  Edscale- Education

5.  Race- White,Black,AIAN,Hawaiian,Asian,Otherrace- converted to binary

6.  Male- female=0,male=1- binary

```{r}
# Checking distribution in the race variable and variable transformation
table(publictransit$White)
publictransit<-publictransit%>%
  mutate(poc=case_when(
           White==1~0,
           White==0~1))
 table(publictransit$poc)

publictransit<-publictransit%>%dplyr::select(-c(White,Black,AIAN,Hawaiian,Asian,Otherrace))

# Merging the 2 datasets (Zipcode and National Community Livability survey)
publictransit<-publictransit%>%
  rename(zipcode=ZipCode)
zip$zipcode<-as.numeric(zip$zipcode)

publictransit<-publictransit%>%
    left_join(zip,by="zipcode")
# Creating the geographic region variable based on the Census definition
publictransit<-publictransit%>%
  mutate(urb_rur=case_when(
    population<2500~"Rural",
    population>=2500 & population<50000~"Urban cluster",
    population>=50000~"Urban"
  ))
# because some zipcodes don't have associated population values- those values are removed
sum(is.na(publictransit$urb_rur))
publictransit<-publictransit %>%
  filter(!is.na(urb_rur))
publictransit<-publictransit%>%
  mutate(type=factor(urb_rur, levels=c("Rural","Urban cluster","Urban")))
```

```{r}
str(publictransit)
# Transforming variables wherever necessary
publictransit$hhincscale<-as.numeric(publictransit$Hhincscale)
publictransit$age<-as.numeric(publictransit$AgeScale)
publictransit$educ<-as.numeric(publictransit$Edscale)
publictransit<-publictransit%>%dplyr::select(-c(Hhincscale,AgeScale,Edscale,zipcode,population,population_density,urb_rur))

```

## Dependent Variables

1.  LifeSat- All things considered, how satisfied are you with your life as a whole these days? (Scale: 0-10)
2.  Comqol- How satisfied are you with the quality-of-life in your community? (Scale: 1-5)

```{r}
# Plotting distributions for the dependent variables
ggplot(publictransit,aes(lifesat))+
  geom_histogram(binwidth=1)+
  geom_text(stat="count",aes(label = ..count..),vjust=0.02)+
  labs(title="Distribution of Life Satisfaction",x="Life Satisfaction")
mean(publictransit$lifesat)
var(publictransit$lifesat)
ggplot(publictransit,aes(qol))+
  geom_bar()+
  geom_text(stat="count",aes(label = ..count..),vjust=0.01)+
  labs(title="Distribution of Satisfaction with Quality of Life in the Community",x=" Satisfaction with Quality of Life" )
```

# Distributions of the Independent Variables

```{r}
table(publictransit$hhveh)
table(publictransit$walkrate)
table(publictransit$ptqual)
```

# Models

## Final Models

```{r}
# Linear regression with interaction effects
lm1<-lm(lifesat~hhveh*ptqual+type*ptqual+walk_fac+ hhincscale+age+educ+poc+male, data=publictransit)
# Ordinal Logistic Regression
logit5<-polr(qol~hhveh*ptqual+type*ptqual+walkrate+ hhincscale+age+educ+poc+male, data=publictransit,method="logistic")
```

### Alternate Models

```{r}
# Linear regression
lm2<-lm(lifesat~car_fac*ptqual+type*ptqual+walkrate+ hhincscale+age+educ+poc+male, data=publictransit)
# Ordinal Logistic Regression
logit2<-polr(qol~car_fac*ptqual+type*ptqual+walk_fac+ hhincscale+age+educ+poc+male, data=publictransit,method="logistic")
# Linear regression for quality of life variable
publictransit$qol_num<-as.numeric(publictransit$qol)
lm_qol<-lm(qol_num~hhveh*ptqual+type*ptqual+walkrate+ hhincscale+age+educ+poc+male, data=publictransit)
```

### Tables for Comparison

```{r}
lmcomp<-stargazer(lm1,lm2, type="text",
          dep.var.labels=c("Satisfaction with Life"),
           covariate.labels=c("# of cars in household (numeric)",
                              " 1 car",
                               "2 cars",
                              "3 cars",
                              "Public Transit Quality (PTQ)-Level 1 (reference:0)",
                               "PTQ Level 2",
                               "PTQ Level 3",
                               "PTQ Level 4",
                                "PTQ Level 5",
                                "Urban Cluster (reference:Rural)",
                                 "Urban(reference:Rural)",
                                 "Walkability Level 2 (reference:1)",
                                  "Walkability Level 3",
                                   "Walkability Level 4",
                                   "Walkability Level 5",
                                    "Walkability Rate (numeric)",
                                    "Household Income",
                                    "Age",
                                  "Education",
                                "Race (0=white,1=person of colour)",
                                   "Gender(0=female,1=male)",
                             "# of cars in household:PTQ Level 1",
                            " # of cars in household:PTQ Level 2",
                            "# of cars in household:PTQ Level 3",
                            "# of cars in household:PTQ Level 4",
                            "# of cars in household:PTQ Level 5",
                            " 1 car: PTQ Level 1", 
                            " 2 cars: PTQ Level 1", 
                            " 3 cars: PTQ Level 1", 
                            " 1 car: PTQ Level 2", 
                            " 2 cars: PTQ Level 2", 
                            " 3 cars: PTQ Level 2",
                            " 1 car: PTQ Level 3", 
                            " 2 cars: PTQ Level 3", 
                            " 3 cars: PTQ Level 3",
                            " 1 car: PTQ Level 4", 
                            " 2 cars: PTQ Level 4", 
                            " 3 cars: PTQ Level 4",
                            " 1 car: PTQ Level 5", 
                            " 2 cars: PTQ Level 5", 
                            " 3 cars: PTQ Level 5",
                            "PTQ Level 1:Urban Cluster",
                            "PTQ Level 2:Urban Cluster",
                            "PTQ Level 3:Urban Cluster",
                            "PTQ Level 4:Urban Cluster",
                            "PTQ Level 5:Urban Cluster",
                            "PTQ Level 1:Urban",
                            "PTQ Level 2:Urban",
                            "PTQ Level 3:Urban",
                            "PTQ Level 4:Urban",
                            "PTQ Level 5:Urban"), single.row=TRUE)
#cat(lmcomp, file = "output_lmcomp.html")

logitcomp1<-stargazer(logit2,logit5,type="text",
          dep.var.labels = c("Satisfaction with Quality of Life"),
          covariate.labels=c(" 1 car",
                               "2 cars",
                              "3 cars",
                             "# of cars in household",
                              "Public Transit Quality (PTQ)-Level 1 (reference:0)",
                               "PTQ Level 2",
                               "PTQ Level 3",
                               "PTQ Level 4",
                                "PTQ Level 5",
                             "Urban Cluster (reference:Rural)",
                                 "Urban(reference:Rural)",
                             "Walkability Level 2 (reference:1)",
                                  "Walkability Level 3",
                                   "Walkability Level 4",
                                   "Walkability Level 5",
                                 "Walkability Rate",
                                    "Household Income",
                                    "Age",
                                  "Education",
                                "Race (0=white,1=person of colour)",
                                   "Gender(0=female,1=male)",
                              " 1 car: PTQ Level 1", 
                            " 2 cars: PTQ Level 1", 
                            " 3 cars: PTQ Level 1", 
                            " 1 car: PTQ Level 2", 
                            " 2 cars: PTQ Level 2", 
                            " 3 cars: PTQ Level 2",
                            " 1 car: PTQ Level 3", 
                            " 2 cars: PTQ Level 3", 
                            " 3 cars: PTQ Level 3",
                            " 1 car: PTQ Level 4", 
                            " 2 cars: PTQ Level 4", 
                            " 3 cars: PTQ Level 4",
                            " 1 car: PTQ Level 5", 
                            " 2 cars: PTQ Level 5", 
                            " 3 cars: PTQ Level 5",
                             "# of cars in household:PTQ Level 1",
                            " # of cars in household:PTQ Level 2",
                            "# of cars in household:PTQ Level 3",
                            "# of cars in household:PTQ Level 4",
                            "# of cars in household:PTQ Level 5",
                            "PTQ Level 1:Urban Cluster",
                            "PTQ Level 2:Urban Cluster",
                            "PTQ Level 3:Urban Cluster",
                            "PTQ Level 4:Urban Cluster",
                            "PTQ Level 5:Urban Cluster",
                            "PTQ Level 1:Urban",
                            "PTQ Level 2:Urban",
                            "PTQ Level 3:Urban",
                            "PTQ Level 4:Urban",
                            "PTQ Level 5:Urban"), single.row=TRUE)
logitcomp2<-stargazer(logit5,lm_qol, type="text",
          dep.var.labels = c("Satisfaction with Quality of Life"),
          covariate.labels=c("# of cars in household",
                              "Public Transit Quality (PTQ)-Level 1 (reference:0)",
                               "PTQ Level 2",
                               "PTQ Level 3",
                               "PTQ Level 4",
                                "PTQ Level 5",
                             "Urban Cluster (reference:Rural)",
                                 "Urban(reference:Rural)",
                                 "Walkability Rate",
                                    "Household Income",
                                    "Age",
                                  "Education",
                                "Race (0=white,1=person of colour)",
                                   "Gender(0=female,1=male)",
                             "# of cars in household:PTQ Level 1",
                            " # of cars in household:PTQ Level 2",
                            "# of cars in household:PTQ Level 3",
                            "# of cars in household:PTQ Level 4",
                            "# of cars in household:PTQ Level 5",
                            "PTQ Level 1:Urban Cluster",
                            "PTQ Level 2:Urban Cluster",
                            "PTQ Level 3:Urban Cluster",
                            "PTQ Level 4:Urban Cluster",
                            "PTQ Level 5:Urban Cluster",
                            "PTQ Level 1:Urban",
                            "PTQ Level 2:Urban",
                            "PTQ Level 3:Urban",
                            "PTQ Level 4:Urban",
                            "PTQ Level 5:Urban"), single.row=TRUE)

#cat(logitcomp, file = "output_logitcomp.html")
# Checking parameters such as AIC, BIC, RMSE
model_performance(logit2)
```

#### used in the paper but does not render properly

logitcomp\<-stargazer(logit2,logit5,lm_qol,type="text", dep.var.labels = c("Satisfaction with Quality of Life","Satisfaction with Quality of Life"), covariate.labels=c(" 1 car", "2 cars", "3 cars", "\# of cars in household", "Public Transit Quality (PTQ)-Level 1 (reference:0)", "PTQ Level 2", "PTQ Level 3", "PTQ Level 4", "PTQ Level 5", "Urban Cluster (reference:Rural)", "Urban(reference:Rural)", "Walkability Level 2 (reference:1)", "Walkability Level 3", "Walkability Level 4", "Walkability Level 5", "Walkability Rate", "Household Income", "Age", "Education", "Race (0=white,1=person of colour)", "Gender(0=female,1=male)", " 1 car: PTQ Level 1", " 2 cars: PTQ Level 1", " 3 cars: PTQ Level 1", " 1 car: PTQ Level 2", " 2 cars: PTQ Level 2", " 3 cars: PTQ Level 2", " 1 car: PTQ Level 3", " 2 cars: PTQ Level 3", " 3 cars: PTQ Level 3", " 1 car: PTQ Level 4", " 2 cars: PTQ Level 4", " 3 cars: PTQ Level 4", " 1 car: PTQ Level 5", " 2 cars: PTQ Level 5", " 3 cars: PTQ Level 5", "\# of cars in household:PTQ Level 1", " \# of cars in household:PTQ Level 2", "\# of cars in household:PTQ Level 3", "\# of cars in household:PTQ Level 4", "\# of cars in household:PTQ Level 5", "PTQ Level 1:Urban Cluster", "PTQ Level 2:Urban Cluster", "PTQ Level 3:Urban Cluster", "PTQ Level 4:Urban Cluster", "PTQ Level 5:Urban Cluster", "PTQ Level 1:Urban", "PTQ Level 2:Urban", "PTQ Level 3:Urban", "PTQ Level 4:Urban", "PTQ Level 5:Urban"), single.row=TRUE)

## Results for the Final Models

## Linear Regression with Interaction Effects

```{r}
lm1_table<-stargazer(lm1,type="text",
          dep.var.labels=c("Satisfaction with Life"),
           covariate.labels=c("# of cars in household",
                              "Public Transit Quality (PTQ)-Level 1 (reference:0)",
                               "PTQ Level 2",
                               "PTQ Level 3",
                               "PTQ Level 4",
                                "PTQ Level 5",
                                "Urban Cluster (reference:Rural)",
                                 "Urban(reference:Rural)",
                                 "Walkability Level 2 (reference:1)",
                                  "Walkability Level 3",
                                   "Walkability Level 4",
                                   "Walkability Level 5",
                                    "Household Income",
                                    "Age",
                                  "Education",
                                "Race (0=white,1=person of colour)",
                                   "Gender(0=female,1=male)",
                             "# of cars in household:PTQ Level 1",
                            " # of cars in household:PTQ Level 2",
                            "# of cars in household:PTQ Level 3",
                            "# of cars in household:PTQ Level 4",
                            "# of cars in household:PTQ Level 5",
                            "PTQ Level 1:Urban Cluster",
                            "PTQ Level 2:Urban Cluster",
                            "PTQ Level 3:Urban Cluster",
                            "PTQ Level 4:Urban Cluster",
                            "PTQ Level 5:Urban Cluster",
                            "PTQ Level 1:Urban",
                            "PTQ Level 2:Urban",
                            "PTQ Level 3:Urban",
                            "PTQ Level 4:Urban",
                            "PTQ Level 5:Urban"), single.row=TRUE)
#cat(lm1_table, file = "output.html")

```

```{r}
# Calculating predicted values for walkability and number of vehicles owned
pred_ols<-ggpredict(lm1, terms="walk_fac")
pred_ols_car<-ggpredict(lm1, terms="hhveh")
#pred_ols
#pred_ols_car
# Graph for Number of Vehicles Owned
ggplot(pred_ols_car, aes(x=x, y=predicted)) +
 geom_point() +
 geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
 linewidth=.3, width=.2,position=position_dodge(.9))+
 labs(x = "Number of Vehicles in the Household", y = "Predicted Scores for Life Satisfaction") +
  ggtitle("Effect of Having Vehicles on Life Satisfaction") 
# Graph for Walkability
ggplot(pred_ols, aes(x=x, y=predicted)) +
 geom_point() +
 geom_errorbar(aes(ymin=conf.low, ymax=conf.high),linewidth=.3, width=.2,position=position_dodge(.9))+
 labs(x = "Perceived Walkability Rate", y = "Predicted Scores for Life Satisfaction",caption="Only significant when walkability rate=5")+
  ggtitle("Perceived Walkability Rate and Life Satisfaction") 
```

## Ordinal Logistic Regression with Interaction Effects

```{r}
logit5table<-stargazer(logit5,type="text",
          dep.var.labels=c("Satisfaction with Quality of Life in the Community"),
           covariate.labels=c("# of cars in household",
                              "Public Transit Quality (PTQ)-Level 1 (reference:0)",
                               "PTQ Level 2",
                               "PTQ Level 3",
                               "PTQ Level 4",
                                "PTQ Level 5",
                                "Urban Cluster (reference:Rural)",
                                 "Urban(reference:Rural)",
                                 "Walkability Rate",
                                    "Household Income",
                                    "Age",
                                  "Education",
                                "Race (0=white,1=person of colour)",
                                   "Gender(0=female,1=male)",
                             "# of cars in household:PTQ Level 1",
                            " # of cars in household:PTQ Level 2",
                            "# of cars in household:PTQ Level 3",
                            "# of cars in household:PTQ Level 4",
                            "# of cars in household:PTQ Level 5",
                            "PTQ Level 1:Urban Cluster",
                            "PTQ Level 2:Urban Cluster",
                            "PTQ Level 3:Urban Cluster",
                            "PTQ Level 4:Urban Cluster",
                            "PTQ Level 5:Urban Cluster",
                            "PTQ Level 1:Urban",
                            "PTQ Level 2:Urban",
                            "PTQ Level 3:Urban",
                            "PTQ Level 4:Urban",
                            "PTQ Level 5:Urban"), single.row=TRUE)

#cat(logit5table, file = "outputlogit5.html")
# Checking parameters such as AIC, BIC, RMSE
model_performance(logit5)

```

## Independent effects- Number of vehicles owned

```{r}
# Calculating predicted probabilities
logitpred_car<-ggpredict(logit5, terms=c("hhveh"))
#logitpred_car
# Selecting only results when quality of life is 1 or 5
filt_car<-logitpred_car%>%filter(response.level%in% c(1,5))
# Graphical representation
ggplot(filt_car, aes(x = x, y = predicted, fill = response.level)) +
      geom_bar(stat = "identity", position = "dodge", width = 0.7) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.4, position = position_dodge(width = 0.7)) +
      theme_minimal(base_size = 13) +
      labs(x = "Number of Vehicles in the Household", y = "Predicted Probability", 
           title = "Predicted Probability for Quality of Life based on Number of Vehicles") +
      labs(fill = "Quality of Life in the Community") +
      scale_fill_manual(values = c("1" = "antiquewhite", 
                                   "5" = "grey39"))+
      theme(legend.position = c(0.05, 0.95), legend.justification = c(0, 1))
```

## Independent effects- Perceived walkability rate

```{r}
# Calculating predicted probabilities
logitpred_walk<-ggpredict(logit5, terms=c("walkrate"))
#logitpred_walk
# Selecting only results when quality of life is 1 or 5
filt<-logitpred_walk%>%filter(response.level%in% c(1,5))
# Graphical representation
ggplot(filt, aes(x = x, y = predicted, fill = response.level)) +
      geom_bar(stat = "identity", position = "dodge", width = 0.7) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.4, position = position_dodge(width = 0.7)) +
      theme_minimal(base_size = 13) +
      labs(x = "Perceived Walkability Rate", y = "Predicted Probability", 
           title = "Predicted Probability for Quality of Life based on \nPerceived Walkability Rate") +
      labs(fill = "Quality of Life in the Community") +
      scale_fill_manual(values = c("1" = "antiquewhite", 
                                   "5" = "grey39"))+
      theme(legend.position = c(0.05, 0.95), legend.justification = c(0, 1))
```

## Independent effects- Perceived public transit quality

```{r}
# Calculating predicted probabilities
logitpred_pt<-ggpredict(logit5, terms=c("ptqual"))
#logitpred_pt
# Selecting only results when quality of life is 1 or 5
filt2<-logitpred_pt%>%filter(response.level%in% c(1,5))
# Graphical representation
ggplot(filt2, aes(x = x, y = predicted, fill = response.level)) +
      geom_bar(stat = "identity", position = "dodge", width = 0.7) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.4, position = position_dodge(width = 0.7)) +
      theme_minimal(base_size = 13) +
      labs(x = "Perceived Public Transit Quality Rate", y = "Predicted Probability", 
           title = "Predicted Probability for Quality of Life based on\nPerceived Public Transit Quality Rate") +
      labs(fill = "Quality of Life in the Community",caption="Significant when Public Transit Quality=5") +
      scale_fill_manual(values = c("1" = "antiquewhite", 
                                   "5" = "grey39"))+
      theme(legend.position = c(0.05, 0.95), legend.justification = c(0, 1))+
  scale_x_discrete(breaks = c(0,1,2,3,4,5), labels = c("No\n public\n transit","1","2","3","4", "Highest quality\n of public\n transit"))
```

## Interaction Effect- Perceived Public Transit Quality Rate and Geographic Region

```{r}
#Calculating predicted probabilites but selecting only 2 levels of perceived public transit quality
pred_logit5_diff<-ggpredict(logit5, terms=c("type", "ptqual[0,5]"))
#pred_logit5_diff
# Selecting only results when quality of life is 1 or 5
predl5_diff <- pred_logit5_diff %>%
    filter(response.level==1|response.level==5)  
#Graphical representation
ggplot(predl5_diff, aes(x = group, y = predicted, fill = response.level)) +
      geom_bar(stat = "identity",  width = 0.7 , position = position_dodge()) +
      facet_grid(. ~ x) +  # Create separate panels for each group
      theme_minimal(base_size = 13) +
      labs(fill="Quality of Life\nin the\nCommunity",x = "Public Transit Quality Rate ", y = "Predicted Probability", 
             title = "Predicted Probability for Quality of Life based on Geographic Area\n and Perceived Public Transit Quality",caption="Significant when Public Transit Quality=5")+
      geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                    linewidth=.3,    # Thinner lines
                    width=.2, position = position_dodge(width=.7))+
      scale_fill_manual(values = c("1" = "antiquewhite", 
                                   "5" = "grey39"))+
  scale_x_discrete(breaks = c(0, 5), labels = c("No\n public\n transit", "Highest quality\n of public\n transit"))
```

## References

Texas A&M Transport Institute. (2017). National Community Livabilty Survey. \[Dataset\]. <https://transit-mobility.tti.tamu.edu/resources/data-from-national-community-livability-survey/>

Ratcliffe, M. (2022, December 22). *Redefining urban areas following the 2020 census*. Census.gov. <https://www.census.gov/newsroom/blogs/random-samplings/2022/12/redefining-urban-areas-following-2020-census.html>

## Note

1.  Used ChatGPT to fix an error with geom_text in the graphs.
2.  Used ChatGPT to figure out how to export stargazer tables.
